Refinements for open automata

### 2022-02-22 2021-Refinement/paper/refinementarticle.tex
proof thm2

For all $x \in X$ we have to consider both  cases:
\begin{align*}
	\text{if } k \in J'_1:\quad &  J'_{23x} \cap H'  
=((J'_{2x}  \setminus \mbrc{k})\uplus J^{2 \prime}_{3} ) \cap (H\uplus J_3 \setminus\{k\} )\\
&=((J'_{2x}\cap (H\uplus J_3))\uplus (J^{2 \prime}_{3}\cap (H\uplus J_3)))\setminus \mbrc{k}\\
&=((J'_{2x}\cap H)\uplus (J^{2 \prime}_{3}\cap  J_3))\setminus \mbrc{k}\text{ since } J_3\cap J'_{2x} = \emptyset \text{ and } H\cap J^{2 \prime}_{3} = \emptyset\\
& =((J'_{2x}\cap H)\uplus J^{2 \prime}_{3})\setminus \mbrc{k}\text{ since } J^{2 \prime}_{3}\subseteq J_3\\
& =(J'_1\cap H)\uplus J^{1 \prime}_{3}\setminus \mbrc{k}\quad \text{ since } J^{1 \prime}_{3}=J^{2 \prime}_{3} \text{ and } J'_1 \cap H = J'_{2x} \cap H\\
&= ((J'_1 \cap H) \uplus (J^{1 \prime}_{3} \cap J_3)) \setminus \mbrc{k}\text{ since } J^{1 \prime}_{3}\subseteq J_3\\ 
& = (J'_1\cap (J_3 \uplus H)) \uplus ((J^{1 \prime}_{3} \cap (J_3 \uplus H)) \setminus \mbrc{k} \text{ since } J_3\cap J'_{1} = \emptyset \text{ and } H\cap J^{1 \prime}_{3} = \emptyset\\
		& = \mpar{(J'_1  \uplus J^{1 \prime}_{3})\setminus \mbrc{k}} \cap \mpar{(J_{3} \uplus H) \setminus \mbrc{k}}\\
		& = J'_{13} \cap H'\\
\text{if } k \notin J'_1:\quad &	J'_{23x} \cap H'  = \mpar{J'_{2x} } \cap \mpar{J_3 \uplus H \setminus \mbrc{k}} \\
& = \mpar{J'_{2x}\cap H } \text{ since } J'_{2x}\cap J_3=\emptyset \land k\not\in J'_{2x} \\
& = \mpar{J'_1\cap H }\\
& = \mpar{J'_{13}\cap H' } \text{ since } J_3\cap J'_1=\emptyset \land k\not\in J_1'\\
\end{align*}
For the rest of the formula we will also consider the two cases separately. 
\begin{itemize}
\item if $k \in J'_1$:\\
Because composition doesn't change the resulting actions, nor their variables, we can extend the valuation context of the variables to cover the variables of the transition $t_3$.   By building  $t_{13}$ and $t_{23}$ we have $\alpha_{13}=
\alpha_{1}$ and $\alpha_{23x}=\alpha_{2x}$. So we deduce: 
$\alpha_{1} = \alpha_{2x} \Rightarrow \alpha_{13} = \alpha_{23x}$.\\
We have also:\\ $\beta_{13j}^{J'_{13}}= \beta_{1j}^{j \in J'_{1}  \setminus \{k\}} \uplus   {(\beta^1_{3j})^{j \in J^{1 \prime}_3} }$
and $\beta_{23xj}^{J'_{23}}= \beta_{2xj}^{j \in J'_{2x}  \setminus \{k\}} \uplus   {(\beta^2_{3j})^{j \in J^{2 \prime}_3} }$\\
So we have:  $\beta_{13j} = \beta_{23xj} \Rightarrow (\beta_{1j} = \beta_{2xj}) \vee (\beta^1_{3j}= \beta^2_{3j})$\\
Considering $\beta^1_{3j}$ and  $\beta^2_{3j}$  are the same (up  renaming) hence we have: $V^1_{3} \uplus V^2_{3}  \vdash {\displaystyle \bigwedge_\subbox{j \in J^{2 \prime}_3} \beta^1_{3j}= \beta^2_{3j}}$\\

Assuming we have $V_{1} \uplus V_{2} \uplus \fvars{t_1}  \vdash {\displaystyle \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj}}$ then we can deduce:
\begin{align*}
& V_{1} \uplus V_{2}  \uplus V^1_{3} \uplus V^2_{3} \uplus \fvars{t_1}\vdash  {\displaystyle \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj}} \vee {\displaystyle \bigwedge_\subbox{j \in J^{2 \prime}_3} \beta^1_{3j}= \beta^2_{3j}}\\
\Rightarrow & V_{13} \uplus V_{23}  \uplus \fvars{t_1}\vdash {\displaystyle \bigwedge_\subbox{j \in (J'_{2x}   \cap H)\uplus J^{2 \prime}_3 } \beta_{13j} = \beta_{23xj}} \\  
\Rightarrow & V_{13} \uplus V_{23}  \uplus \fvars{t_1}\vdash {\displaystyle \bigwedge_\subbox{j \in \Rightarrow   ((J'_{2x}   \cap H)\uplus J^{2 \prime}_3)\setminus \{k\} } \beta_{13j} = \beta_{23xj}} \\
\Rightarrow &  V_{13} \uplus V_{23}  \uplus \fvars{t_1}\vdash {\displaystyle \bigwedge_\subbox{j \in (J'_{23x}   \cap H') } \beta_{13j} = \beta_{23xj}}\\
\end{align*}
On another side,   the composition gives $g_{13} \Leftrightarrow g_1 \wedge g^1_3 \wedge \alpha^1_3=\beta_{1k}$ and
$g_{23x}  \Leftrightarrow g_{2x} \wedge g^2_3 \wedge \alpha^2_3=\beta_{2xk}$. \\
In addition as $k \in H$ we have   
$\beta_{1k} = \beta_{2xk}$ then we  deduce:
$g^1_3 \wedge \alpha^1_3=\beta_{1k}  \Leftrightarrow g^2_3 \wedge \alpha^2_3= \beta_{2xk}$

\TODO{The subtlety here is that the original relation implies that valuations in A3 are equal (ie the value for each variable are the same in both valuations, modulo renaming), after a transition we should obtain ``equal'' valuations because Post are deterministic}

@@

By hypothesis we have  \(A_1\subst{A_3}{k}\) is non-blocking,  then according to the Definition \ref{Def:Reach} we have $ \sigma_{013} \vdash \reach{A_{13}}\mpar{s_{013}}$ and for all $t_{13} \in T_{13}$:
\[ \fvars{t_{13}} \vdash \left(\reach{A_{13}}\mpar{s_{13}} \wedge g_{13} \implies \reach{A_{13}}\mpar{s_{13}'}\psubst{\psi_{13}}\right) \]	
From these two previous formulas we obtain:
\begin{multline*}
V_{13} \uplus V_{23} \uplus \fvars{t_1}  \uplus \fvars{t_{13}} \vdash\\
R\mpar{s_1, s_{2}} \wedge g_{13}  \wedge \reach{A_{13}}\mpar{s_{13}}
%\wedge \bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3\wedge s^1_3=s^2_3
\implies  \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_{13} = \alpha_{23x} \wedge \bigwedge_\subbox{j \in J'_{23x} \cap H'} \beta_{13j} = \beta_{23xj} \\
			\nwedge g_{23x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}} 
%\wedge \bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3\wedge s^1_3=s^2_3
		\end{array}}   \wedge \reach{A_{13}}\mpar{s_{13}'}\psubst{\psi_{13}} 
\end{multline*}

Considering that ate each step the valuations $\sigma_{3}^1$ and $\sigma_{3}^2$ associate the same values to the ``same'' variables modulo renaming, we have the tautology:\\ $\left({\displaystyle \bigwedge_\subbox{v_3\in V_3} v^1_3=v^2_3\wedge s^1_3 = s^2_3}\right)\psubst{\sigma_{3}^1 \uplus \sigma_{3}^2 }$.

\end{itemize}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%555

\TODO{Pas termine: valider d'abord la definition de R}
\item The last formula to prove is the following:
\begin{multline*}
\forall \mpar{s_{13}, s_{23}} \in S_{13} \times S_{23},\\ V_{13} \uplus V_{23} \uplus \biguplus_\subbox{t_{23}\in \fOT{s_{23}}} \fvars{t_2} \vdash \left(R\mpar{s_{13}, s_{23}} \wedge \bigvee_\subbox{t_{23} \in \fOT{s_{23}}} \fguard{t_{23}} \implies \bigvee_\subbox{t_1 \in \fOT{s_{13}}} \fguard{t_{13}} \right)
\end{multline*}
By hypothesis we have the following:
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2,\\ V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash \left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
\end{multline*}
By adding the term $\displaystyle{\bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3}}$ to both sides of the implication, we get: 
\begin{multline*}
\forall \mpar{s_1, s_2,s_3} \in S_1 \times S_2 \times S_3, V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash\\ \left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3}\right)
\end{multline*}
By hypothesis we also have \(A_1\subst{A_3}{k}\) is non-blocking,  thus we can write the following:
\[ \fvars{t_{13}} \vdash \left(\reach{A_{13}}\mpar{s_{13}} \wedge g_{13} \implies \reach{A_{13}}\mpar{s_{13}'}\psubst{\psi_{13}}\right) \]	
\end{enumerate}

>>>>

For all $x \in X$ we have to consider both  cases:
	\text{if } k \in J'_1:\quad &  J'_{23x} \cap H'  
=((J'_{2x}  \setminus \mbrc{k})\uplus J^{2 \prime}_{3} ) \cap (H\uplus J_3 \setminus\{k\} )\\
&=((J'_{2x}\cap (H\uplus J_3))\uplus (J^{2 \prime}_{3}\cap (H\uplus J_3)))\setminus \mbrc{k}\\
&=((J'_{2x}\cap H)\uplus (J^{2 \prime}_{3}\cap  J_3))\setminus \mbrc{k}\text{ since } J_3\cap J'_{2x} = \emptyset \text{ and } H\cap J^{2 \prime}_{3} = \emptyset\\
& =((J'_{2x}\cap H)\uplus J^{2 \prime}_{3})\setminus \mbrc{k}\text{ since } J^{2 \prime}_{3}\subseteq J_3\\
& =(J'_1\cap H)\uplus J^{1 \prime}_{3}\setminus \mbrc{k}\quad \text{ since } J^{1 \prime}_{3}=J^{2 \prime}_{3} \text{ and } J'_1 \cap H = J'_{2x} \cap H\\
&= ((J'_1 \cap H) \uplus (J^{1 \prime}_{3} \cap J_3)) \setminus \mbrc{k}\text{ since } J^{1 \prime}_{3}\subseteq J_3\\ 
& = (J'_1\cap (J_3 \uplus H)) \uplus ((J^{1 \prime}_{3} \cap (J_3 \uplus H)) \setminus \mbrc{k} \text{ since } J_3\cap J'_{1} = \emptyset \text{ and } H\cap J^{1 \prime}_{3} = \emptyset\\
		& = \mpar{(J'_1  \uplus J^{1 \prime}_{3})\setminus \mbrc{k}} \cap \mpar{(J_{3} \uplus H) \setminus \mbrc{k}}\\
		& = J'_{13} \cap H'\\
\text{if } k \notin J'_1:\quad &	J'_{23x} \cap H'  = \mpar{J'_{2x} } \cap \mpar{J_3 \uplus H \setminus \mbrc{k}} \\
& = \mpar{J'_{2x}\cap H } \text{ since } J'_{2x}\cap J_3=\emptyset \land k\not\in J'_{2x} \\
& = \mpar{J'_1\cap H }\\
& = \mpar{J'_{13}\cap H' } \text{ since } J_3\cap J'_1=\emptyset \land k\not\in J_1'\\
For the rest of the formula we will also consider the two cases separately. 
\begin{itemize}
\item if $k \in J'_1$:\\
The proof is based on the following facts.
First, because composition doesn't change the resulting actions, nor their variables, we can extend the valuation context of the variables to cover the variables of the transition $t_3$.   By building  $t_{13}$ and $t_{23}$ we have $\alpha_{13}=
\alpha_{1}$ and $\alpha_{23x}=\alpha_{2x}$. So we deduce: 
$\alpha_{1} = \alpha_{2x} \Rightarrow \alpha_{13} = \alpha_{23x}$.\\
Second, we have:\\ $\beta_{13j}^{J'_{13}}= \beta_{1j}^{j \in J'_{1}  \setminus \{k\}} \uplus   {(\beta^1_{3j})^{j \in J^{1 \prime}_3} }$
and $\beta_{23xj}^{J'_{23}}= \beta_{2xj}^{j \in J'_{2x}  \setminus \{k\}} \uplus   {(\beta^2_{3j})^{j \in J^{2 \prime}_3} }$\\
Therefore, we have:  $\beta_{13j} = \beta_{23xj} \Rightarrow (\beta_{1j} = \beta_{2xj}) \vee (\beta^1_{3j}= \beta^2_{3j})$\\
Third, considering $\beta^1_{3j}$ and  $\beta^2_{3j}$  are the same (up  renaming) hence we have: $V^1_{3} \uplus V^2_{3}  \vdash {\displaystyle \bigwedge_\subbox{j \in J^{2 \prime}_3} \beta^1_{3j}= \beta^2_{3j}}$\\
And suppose that  $V_{1} \uplus V_{2} \uplus \fvars{t_1}  \vdash {\displaystyle \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj}}$ holds then we can deduce:
& V_{1} \uplus V_{2}  \uplus V^1_{3} \uplus V^2_{3} \uplus \fvars{t_1}\vdash  & {\displaystyle \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj}} \vee {\displaystyle \bigwedge_\subbox{j \in J^{2 \prime}_3} \beta^1_{3j}= \beta^2_{3j}}\\
\Rightarrow & V_{13} \uplus V_{23}  \uplus \fvars{t_1}\vdash & {\displaystyle \bigwedge_\subbox{j \in (J'_{2x}   \cap H)\uplus J^{2 \prime}_3 } \beta_{13j} = \beta_{23xj}} \\  
\Rightarrow & V_{13} \uplus V_{23}  \uplus \fvars{t_1}\vdash & {\displaystyle \bigwedge_\subbox{j \in \Rightarrow   ((J'_{2x}   \cap H)\uplus J^{2 \prime}_3)\setminus \{k\} } \beta_{13j} = \beta_{23xj}} \\
\Rightarrow &  V_{13} \uplus V_{23}  \uplus \fvars{t_1}\vdash & {\displaystyle \bigwedge_\subbox{j \in (J'_{23x}   \cap H') } \beta_{13j} = \beta_{23xj}}\\
Finally, the composition gives $g_{13} \Leftrightarrow g_1 \wedge g^1_3 \wedge \alpha^1_3=\beta_{1k}$ and
$g_{23x}  \Leftrightarrow g_{2x} \wedge g^2_3 \wedge \alpha^2_3=\beta_{2xk}$. \\
Furthermore, as $k \in H$ we have   
$\beta_{1k} = \beta_{2xk}$ then we  deduce:
$g^1_3 \wedge \alpha^1_3=\beta_{1k}  \Leftrightarrow g^2_3 \wedge \alpha^2_3= \beta_{2xk}$

@@

By hypothesis we have  \(A_1\subst{A_3}{k}\) is non-blocking,  then according to the Definition \ref{Def:Reach} we have $ \sigma_{013} \vdash \reach{A_{13}}\mpar{s_{013}}$ and for all $t_{13} \in T_{13}$:
\[ \fvars{t_{13}} \vdash \left(\reach{A_{13}}\mpar{s_{13}} \wedge g_{13} \implies \reach{A_{13}}\mpar{s_{13}'}\psubst{\psi_{13}}\right) \]	
From these two previous formulas we obtain:
\begin{multline*}
V_{13} \uplus V_{23} \uplus \fvars{t_1}  \cap \fvars{t_{13}} \vdash\\
R\mpar{s_1, s_{2}} \wedge g_{13}  \wedge \reach{A_{13}}\mpar{s_{13}}
%\wedge \bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3\wedge s^1_3=s^2_3
\implies  \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_{13} = \alpha_{23x} \wedge \bigwedge_\subbox{j \in J'_{23x} \cap H'} \beta_{13j} = \beta_{23xj} \\
			\nwedge g_{23x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}} 
%\wedge \bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3\wedge s^1_3=s^2_3
		\end{array}}   \wedge \reach{A_{13}}\mpar{s_{13}'}\psubst{\psi_{13}} 
\end{multline*}	
\TODO{The subtlety here is that the original relation implies that valuations in A3 are equal (ie the value for each variable are the same in both valuations, modulo renaming), after a transition we should obtain ``equal'' valuations because Post are deterministic}

Considering that at each step the valuations $\sigma_{3}^1$ and $\sigma_{3}^2$ associate the same values to the ``same'' variables modulo renaming, so for any states $s^1_3$ and $s^2_3$ the following formula is a tautology:\\ $\left({\displaystyle \bigwedge_\subbox{v_3\in V_3} v^1_3=v^2_3\wedge s^1_3 = s^2_3}\right)\psubst{\sigma_{3}^1 \uplus \sigma_{3}^2 }$ which is equivalent to \\
$V_3^1 \uplus V_3^2 \vdash \left({\displaystyle \bigwedge_\subbox{v_3\in V_3} v^1_3=v^2_3\wedge s^1_3 = s^2_3}\right)$ \\
By adding tautologies to both members of the implication, we get:
\begin{multline*}
V_{13} \uplus V_{23} \uplus \fvars{t_1}  \cap \fvars{t_{13}} \vdash\\
R\mpar{s_1, s_{2}} \wedge g_{13}  \wedge \reach{A_{13}}\mpar{s_{13}}
\wedge \bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3\wedge s^1_3=s^2_3
\implies \hspace{5em} \\ \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_{13} = \alpha_{23x} \wedge \bigwedge_\subbox{j \in J'_{23x} \cap H'} \beta_{13j} = \beta_{23xj} \nwedge \\
			g_{23x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}} \wedge \reach{A_{13}}\mpar{s_{13}'}\psubst{\psi_{13}} 
\wedge \bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3\wedge s^{1\prime}_3=s^{2\prime}_3
		\end{array}}   
\end{multline*}	
\TODO{Point a verifier!}
\item if $k \notin J'_1$:\\
The first and second arguments remain the same as the previous case, while for the third we have: $g_{13} \Leftrightarrow g_1$ and
$g_{23x}  \Leftrightarrow g_{2x} $. By replacing  these results in the formula $(*)$, we get the same formula:
\begin{multline*}
V_{13} \uplus V_{23} \uplus \fvars{t_1}  \vdash\\
R\mpar{s_1, s_{2}} \wedge g_{13}  
%\wedge \bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3\wedge s^1_3=s^2_3
\implies  \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_{13} = \alpha_{23x} \wedge \bigwedge_\subbox{j \in J'_{23x} \cap H'} \beta_{13j} = \beta_{23xj} \\
			\nwedge g_{23x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
%\wedge \bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3\wedge s^1_3=s^2_3
		\end{array}}   
\end{multline*}	
From this point on, the proof continues as in the previous case.\\
\end{itemize}

\item The last formula we need to prove is the following: for all $\mpar{s_{13}, s_{23}} \in S_{13} \times S_{23}$
\begin{multline*}
V_{13} \uplus V_{23} \uplus \biguplus_\subbox{t_{23}\in \fOT{s_{23}}} \fvars{t_{23}} \vdash \left(R^{\prime}\mpar{s_{13}, s_{23}} \wedge \bigvee_\subbox{t_{23} \in \fOT{s_{23}}} \fguard{t_{23}} \implies \bigvee_\subbox{t_1 \in \fOT{s_{13}}} \fguard{t_{13}} \right)
\end{multline*}
Let's start  with the hypothesis of the theorem that $\wrel{A_1}{A_2}{H}$, thus we have for all  $\mpar{s_1, s_2} \in S_1 \times S_2$ 
\begin{multline*}
V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash \left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
\end{multline*}
By adding  the same term to both members of the implication, we get:
\begin{multline*}
V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash \\\left(R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \wedge  \reach{A_{13}}\mpar{s_{13}}\implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}  \wedge \reach{A_{13}}\mpar{s_{13}}\right)
\end{multline*}
But by hypothesis we have also the result of the composition \(A_1\subst{A_3}{k}\) is non-blocking, which leads to:
\begin{multline*}
V_{13}  \uplus \biguplus_\subbox{t_1 \in \fOT{s_1}} \fvars{t_1} \vdash \left( \reach{A_{13}}\mpar{s_{13}} \wedge \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \implies \bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}} \right)
\end{multline*}
